---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# distcrete

[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![](http://cranlogs.r-pkg.org/badges/distcrete)](http://cran.rstudio.com/packages=distcrete)


[![Travis-CI Build Status](https://travis-ci.org/reconhub/distcrete.svg?branch=master)](https://travis-ci.org/reconhub/distcrete)
[![Build status](https://ci.appveyor.com/api/projects/status/uy5out71p9qkh93i/branch/master?svg=true)](https://ci.appveyor.com/project/thibautjombart/distcrete/branch/master)
[![codecov.io](https://codecov.io/github/reconhub/distcrete/coverage.svg?branch=master)](https://codecov.io/github/reconhub/distcrete?branch=master)


> Discrete Distribution Approximations

`distcrete` takes a distribution and a set of parameters and returns a list object with functions as elements. Each one is the equivalent to the function calls we typically expect to be able to do in R for a given distribution: 

- Density e.g.`dnorm` 
- Distribution function e.g. `pnorm`
- Quantile function e.g. `qnorm`
- Random generation e.g. `rnorm`

Each function created by `distcrete` corresponds to the first letter of the non-discrete equivalent.

```{r example}
set.seed(415)
d0 <- distcrete::distcrete("gamma", 1, shape = 3, w = 0)
d0$d(1:10)
d0$p(c(.1,.5))
d0$q(c(.1,.5))
d0$r(10)
```

## Allowed distributions
You can use any distribution that conforms to the following expectations:

- It has a distribution function like `p[dist name]` available
- It has a quantile function like `q[dist name]` available

These can be loaded from a package or created on the fly, but must exist when the `distcrete()` function is called.

## Installation

You can install it from CRAN with:
```{r cran-installation, eval = FALSE}
install.packages("distcrete")
```

You can install `distcrete` the most up to date version from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("reconhub/distcrete")
```

## Tests
```{r tests}
devtools::test()
```
