---
title: "ML estimation with distcrete"
author: "Thibaut Jombart"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{distcrete_ML}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

``` {r echo = FALSE, results = "hide"}
knitr::opts_chunk$set(
  error = FALSE,
  fig.width = 7,
  fig.height = 5)
set.seed(1)
```

In this vignette, we illustrate how *distcrete* can be combined with classical optimisation procedures available in R to derive maximum-likelihood (ML) of parameters of distributions of interest. We simulate simple data from a discretised exponential distribution, and then attempt to recover the original parameters using ML.


## Simulating data

Simulating data from a discretised distribution is very easy using *distcrete*, as discretised distributions already contain a random variate generator in the `$r` component. We illustrate this by generating a discretised exponential distribution `x` with rate parameter 0.123:
```{r}
library(distcrete)
rate <- 0.123
plot(function(x) dexp(x, rate), xlim = c(0, 60),
     main = "Original Exponential distribution")
x <- distcrete("exp", interval = 1L, rate)
x$r
```

We simulate 200 draws from the distribution `x`:
```{r, simulation}
set.seed(1)
sim <- x$r(200)
head(sim)
summary(sim)
plot(table(sim), lwd=3, xlab = "x", ylab = "Frequency",
     main = "Simulated sample")
```



## ML estimation

We will use the base function `optim` to find maximum likelihood estimates of the shape and the rate of the distribution `x`. The log-likelihood function to be optimised can be written as:
```{r, target}
ll <- function(param) {
   d <- distcrete("exp", interval = 1L, param)$d
   sum(d(sim, log = TRUE))
}
```

For instance, the log-likelihood with shapes and rates of 1 can be computed as:
```{r, ll_example}
param <- c(rate = 1)
ll(param)
```


We can feed this function and initial parameters to `optim`:
```{r, optimise}
opt <- optimise(ll, c(0, 20), maximum = TRUE)
opt
```

The ML estimate for the rate of the distribution is `r opt$maximum`, which is close enough to the original value of `r rate`.




## Discretised gamma example

We repeat the same exercise with a gamma distribution. The only difference with the previous example is that `optim` needs to be used, as the parameter space now has 2 dimensions.

We start by simulating a sample from a discretised gamma distribution:
```{r, gamma}
shape <- 2.5
rate <- 0.6
y <- distcrete("gamma", interval = 1L, shape, rate)
set.seed(1)
sim2 <- y$r(200)

head(sim2)
summary(sim2)
plot(table(sim2), lwd=3, xlab = "x", ylab = "Frequency",
     main = "Simulated sample")

```

We create a log-likelihood function for these data, as well as a deviance (it is easier to minimise the deviance with `optim`):
```{r, ll_gamma}
ll2 <- function(param) {
   d <- distcrete("gamma", interval = 1L, param[1], param[2])$d
   sum(d(sim2, log = TRUE))
}

dev2 <- function(param) -2 * ll2(param)

optim(c(1,1), dev2)
```

We can verify that the estimates are not dependent on the initial state:
```{r, verif}
optim(c(0.5,20), dev2)
```